---
title: "Blog 1: Switzerland's COVID-19 numbers"
description: |
  A look into the coronavirus statistics in Switzerland.
author:
  - name: Justin Thomas 
    affiliation: ETC5523, Monash University
    affiliation_url: https://cwd.numbat.space/
date: 09-03-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(gridExtra)
library(plotly)
library(readr)
library(DT)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

# Prelude

Switzerland has been battling the Coronavirus ever since it rocked the beautiful moutaineous region in the EU. In this blog, we take a further look into the magnitude of the country's Covid-19 numbers.

# Data Description

```{r}
deaths <- read_csv("deaths.csv")
```

```{r}
countries <- read_csv("country-list.csv")
```

```{r}
infections <- read_csv("infections.csv")
```

```{r}
death_case <- read_csv("deaths-cases.csv")
```

The data used in this blog contains records for the number of new daily cases and deaths as well as the cumulative number of deaths and cases from February 26 to August 29. 

## Data source

*Coronavirus: the situation in Switzerland*. Retrieved August 30, 2020, from https://www.swissinfo.ch/eng/covid-19_coronavirus--the-situation-in-switzerland/45592192


# Analysis

Table \@ref(fig:deaths-cases) will help you visualize the total deaths and cases. It is also interactive, so adjust the dates to get a better understanding of how the numbers from changing across different time periods.

```{r deaths-cases, fig.cap="Cumulative deaths and cases from February 26 to August 28"}
death_case <- death_case %>% rename(Date = date, 
                      `Cumulative deaths` = Deaths,
                      `Cumulative cases` = Cases)

death_case$Date <- as.Date(death_case$Date)
      
death_case %>% datatable(filter = 'top') %>% 
        formatStyle('Cumulative deaths', color = 'black', backgroundColor = 'yellow') %>%          formatStyle('Cumulative cases', color = 'black', backgroundColor = 'yellow') %>%           formatStyle('Date', color = 'white', backgroundColor = 'green')
```


Table \@ref(fig:summary-statistic) below displays the mean, minimum and maximum values for the daily new cases and deaths.

```{r summary-statistic, fig.cap="Summary statistics of daily new deaths and infection cases"}
infections <- infections %>% rename(Date = date,
                      `New infections` = new.infections)

infections$Date <- as.Date(infections$Date)

summaryi <- infections %>% summarise(Mean = round(mean(`New infections`), digits = 0), Min = min(`New infections`), Max = max(`New infections`))

summaryd <- deaths %>% summarise(Mean = round(mean(`New deaths`), digits = 0), Min = min(`New deaths`), Max = max(`New deaths`))

infection_deaths <- bind_rows(summaryd, summaryi)

rownames(infection_deaths) <- c("Daily deaths", "Daily infections")

infection_deaths %>% datatable() %>% 
    formatStyle('Mean', color = 'black', backgroundColor = 'yellow') %>% 
    formatStyle('Min', color = 'black', backgroundColor = 'yellow') %>% 
    formatStyle('Max', color = 'black', backgroundColor = 'yellow') 
```


Although new cases are being recorded at an average of 223 people a day, it is interesting to note that roughly 5% of all infected individuals die every day. Something more interesting about this is that...stay tuned for the next blog!



