---
title: "Blog 2: Singapore's COVID-19 numbers"
description: |
  A look into the coronavirus statistics in Singapore.
author:
  - name: Justin Thomas 
    affiliation: ETC5523, Monash University
    affiliation_url: https://cwd.numbat.space/
date: 09-17-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(gridExtra)
library(plotly)
library(readr)
library(DT)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

# Prelude

Singapore has received praise in its response to managing the Coronavirus pandemic. The small nation in South East Asia has started to see a dramatic decrease in cases over the recent weeks. In this blog, we take a further look into the trend of the country's COVID-19 cases, deaths and hospitalization numbers.

# Data Description

```{r}
figures <- read_csv("covid_19_sg.csv")
figures$date <- as.Date(figures$date)
```

The data used in this blog contains the following records from January 23 to September 12:

- new daily cases
- new daily deaths
- cumulative deaths 
- cumulative cases
- daily discharged from hospital
- cumulative discharged from hospital
- ICU cases
- general ward cases
- discharged to isolation
- isolation cases
- cumulative cases who have completed isolation
- daily imported cases
- daily local transmission cases


The data is sourced from:

Chua, H. (2020, September 16). *COVID-19 Singapore*. Retrieved September 12, 2020, from https://data.world/hxchua/covid-19-singapore


# Analysis

Table \@ref(fig:deaths-cases) will help you visualize the total deaths and cases. It is also interactive, so adjust the dates to get a better understanding of how the numbers change across different time periods.

```{r deaths-cases, eval=FALSE, fig.cap="Cumulative deaths and cases from February 26 to August 28"}
death_case <- death_case %>% rename(Date = date, 
                      `Cumulative deaths` = Deaths,
                      `Cumulative cases` = Cases)

death_case$Date <- as.Date(death_case$Date)
      
death_case %>% datatable(filter = 'top') %>% 
        formatStyle('Cumulative deaths', color = 'black', backgroundColor = 'yellow') %>%          formatStyle('Cumulative cases', color = 'black', backgroundColor = 'yellow') %>%           formatStyle('Date', color = 'white', backgroundColor = 'green')
```


Table \@ref(fig:summary-statistic) below displays the mean, minimum and maximum values for the daily new cases and deaths.

```{r summary-statistic, eval=FALSE, fig.cap="Summary statistics of daily new deaths and infection cases"}
infections <- infections %>% rename(Date = date,
                      `New infections` = new.infections)

infections$Date <- as.Date(infections$Date)

summaryi <- infections %>% summarise(Mean = round(mean(`New infections`), digits = 0), Min = min(`New infections`), Max = max(`New infections`))

summaryd <- deaths %>% summarise(Mean = round(mean(`New deaths`), digits = 0), Min = min(`New deaths`), Max = max(`New deaths`))

infection_deaths <- bind_rows(summaryd, summaryi)

rownames(infection_deaths) <- c("Daily deaths", "Daily infections")

infection_deaths %>% datatable() %>% 
    formatStyle('Mean', color = 'black', backgroundColor = 'yellow') %>% 
    formatStyle('Min', color = 'black', backgroundColor = 'yellow') %>% 
    formatStyle('Max', color = 'black', backgroundColor = 'yellow') 
```


Although new cases are being recorded at an average of 223 people a day, it is interesting to note that on average, roughly 5% of all infected individuals die every day. 5% may seem like a lot for average daily deaths, but this is about 11 deaths on average every day. When you consider this against other major European countries like France, Spain and Italy, it is far less and this low figure is a result of Switzerland's early response to the pandemic. Clare O'Dea (2020) reports that because of an early lockdown and an extensive testing, tracing and quarantine regime, the feared disaster of a traumatic death toll and overwhelmed hospitals never came to pass. 

Furthermore, the fact that all Swiss residents were always free to go outside for exercise without any permit system like those imposed in France or Italy (Clare O'Dea, 2020) without significantly increasing case numbers and the death toll is a testament to how effective their safety measures were during the early phase of the pandemic.  

Switzerland's proactive approach to getting testing done as soon as possible whilst also ensuring their people are given freedom of movement within the country and around EU has been well received by the citizens. This is perhaps why Switzerland was ranked the safest country in the world in June 2020 by Hong-Kong based Deep Knowledge Group (Clare O'Dea, 2020).

# References

Chua, H. (2020, September 16). *COVID-19 Singapore*. Retrieved September 12, 2020, from https://data.world/hxchua/covid-19-singapore

